ms2

### 2017-08-16 ms2/ms2_display.Rmd
ms2 fewer plots

```{r load_data}
ineq = read_tsv('ineq.tsv') %>%
  group_by(drug_group, unit_type, unit) %>%
  summarize_at(vars(mean, fnz), mean) %>%
  ungroup()

@@

```{r hrr_linear}
results %>%
  filter(model=='univariate', unit_type=='hrr', term=='fnz') %>%
  interval_plot() +
  geom_hline(yintercept=1, col='gray') +
  ylab(expression(beta[f])) +
  ggtitle("Linear univariate, HRR")
```

```{r state_linear}
results %>%
  filter(model=='univariate', unit_type=='state', term=='fnz') %>%
  interval_plot() +
  ylab(expression(beta[f])) +
  ggtitle("Linear univariate, state")
```

@@

- *E. faecium* and $\beta$-lactams have a postive relationship (implying higher mean consumption, given the same fraction of people consuming, goes with more resistance)
- *E. coli* and Bactrim have a *negative* relationship (higher consumption goes with *less* resistance)
- *S. pneumoniae* and tetracyclines have a negative relationship

@@

- *E. coli* and $\beta$-lactams, negative relationship, HRR only
- Non-MRSA *S. aureus* and first-generation cephalosporins, negative, HRR only
- *E. faecium* and nitro, negative, both HRR and state
- *S. aureus* and clinda, positive, state only
- *S. aureus* and tetracyclines, positive, state only
- *E. cloacae* and quinolones, positive, state only

The positive relationships are not troubling: it means that it might be that increasing consumption among the consumers can increase resistance. The negative relationships are troubling: how could decreasing consumption among consumers increase resistance?

### E. faecium and beta-lactams

This one doesn't worry me. First, the univariate regression $R \sim f_>$ isn't significant. Second, the ANOVA gives different results depending on the order of the variables: if $\mu$ goes first, it is non-significant but $f_>$, going second, is; conversely, if $f_>$ goes first, then $\mu$ is significant. So I think this is a weird case about collinearity.

```{r efaecium_betalactams}
ineq %>%
  filter(unit_type=='hrr') %>%
  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='Ec. faecium', drug_group=='beta_lactam') %>%
  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed beta-lactams') +
  ylab('Beta-lactam resistance E. faecium') +
  ggtitle('E. faecium & beta-lactams across HRRs')

ineq %>%
  filter(unit_type=='state') %>%
  rename(state=unit) %>%
  left_join(regions, by='state') %>%
  right_join(state_abg, by=c('state', 'drug_group')) %>%
  filter(bug=='Ec. faecium', drug_group=='beta_lactam') %>%
  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed beta-lactams') +
  ylab('Beta-lactam resistance E. faecium') +
  ggtitle('E. faecium & beta-lactams across states')
```

@@

From the Spearman and univariate models, Bactrim resistance in *E. coli* is correlated with increased resistance (black line). When breaking that apart by region, however, there are different relationship (colored lines).

@@

### S. pneumonia and tetracyclines

This one doesn't bother me either. Tetracycline resistance in pneumococcus not significantly related ot tetracycline consumption in the univariate models; the coefficient is almost exactly zero. If you run a regression $R \sim \mu$, you get a non-significant result. So you only get this relationship *after* you correct for an approximately flat relationship. Not promising.

```{r pneumo_tetra}
ineq %>%
  filter(unit_type=='hrr') %>%
  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='S. pneumoniae', drug_group=='tetracycline') %>%
  filter(mean_percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed Bactrim') +
  ylab('Bactrim-resistance E. coli') +
  ggtitle('S. pneumonia & Bactrim across HRRs')
```

>>>>

```{r load_data, cache.extra=c(md5sum('ineq.tsv'), md5sum('results.tsv'))}
ineq = read_tsv('ineq.tsv') %>%
  group_by(drug_group, unit_type, unit) %>%
  summarize_at(vars(mean, fnz), mean) %>%
  ungroup()

@@

```{r hrr_linear}
results %>%
  filter(model=='univariate_fnz', unit_type=='hrr', term=='fnz') %>%
  interval_plot() +
  geom_hline(yintercept=1, col='gray') +
  ylab(expression(beta[f])) +
  ggtitle("Linear univariate, HRR")
```

```{r}
results %>%
  filter(model=='univariate_mean', unit_type=='hrr', term=='mean') %>%
  interval_plot() +
  geom_hline(yintercept=1, col='gray') +
  ylab(expression(beta[f])) +
  ggtitle("Linear univariate, HRR")
```

```{r state_linear}
results %>%
  filter(model=='univariate_fnz', unit_type=='state', term=='fnz') %>%
  interval_plot() +
  ylab(expression(beta[f])) +
  ggtitle("Linear univariate, state")
```

@@

- *E. coli* and Bactrim have a *negative* relationship (higher consumption goes with *less* resistance)
- *E. faecium* and $\beta$-lactams have a postive relationship (implying higher mean consumption, given the same fraction of people consuming, goes with more resistance)
- *S. pneumoniae* and tetracyclines have a negative relationship

The second two, *E. faecium* with $\beta$-lactams and pneumococcus with tetracyclines, are not problematic: in both cases, neither $R \sim f_>$ nor $R \sim \mu$ is significant. We're only interested in cases where $f_>$ gives information and $\mu$ gives information beyond that. There are cases where $f_>$ doesn't appear to give any information.

@@

- *E. coli* and $\beta$-lactams, negative relationship, HRR only; significant in univariate regressions $R \sim f_>$
- *E. cloacae* and quinolones, positive, state only; marginally significant in univariate
- Non-MRSA *S. aureus* and first-generation cephalosporins, negative, HRR only; but not significant in univariate
- *E. faecium* and nitro, negative, both HRR and state; not significant in univariate
- *S. aureus* and clinda, positive, state only; not significant in univariate
- *S. aureus* and tetracyclines, positive, state only; not significant in univariate

@@

This negative relationship is the most tricky.

First, the regression results are equivocal. Regressing $r ~ \mu$, where $r$ are the residuals from the regression $R ~ f_>$, gives a non-significant result. An ANOVA analysis shows that sequential application of $f_>$ and $\mu$ as predictors of $R$ both explain non-trivial variance; but if the order is reversed, and $\mu$ is used first, then it *doesn't* explain any variance. This makes me think that there is just some weird interaction between the two terms in the multiple regression.

Second, from the Spearman and univariate models, Bactrim resistance in *E. coli* is correlated with increased resistance (black line). When breaking that apart by region, however, there are different relationship (colored lines). This makes it seem like there is something strange going on.

@@

### E. coli and beta-lactams

This negative relationship was significant only for HRRs, not states. Not surprisingly, there is something weird going on with the HRRs: the Southern HRRs are all above the $R \sim f_>$ line, suggesting that the relationship with the mean is just a predictor about region.

```{r ecoli_betalactam}
ineq %>%
  filter(unit_type=='hrr') %>%
  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='beta_lactam') %>%
  filter(mean_percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed beta-lactams') +
  ylab('Beta-lactam resistance in E. coli') +
  ggtitle('E. coli & beta-lactams across HRRs')

ineq %>%
  filter(unit_type=='state') %>%
  rename(state=unit) %>%
  left_join(regions, by='state') %>%
  right_join(state_abg, by=c('state', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='beta_lactam') %>%
  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed beta-lactams') +
  ylab('Beta-lactam resistance E. coli') +
  ggtitle('E. coli & beta-lactams across states')
```
